---
title: 'Interactive Visualizations'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align='center', warning=F, message=F)
```

```{r include=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(tidyr)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(ggcorrplot)
library(stringr)
library(knitr)
library(kableExtra)
library(plotly)
library(widgetframe)
library(withr)
library(treemapify)
options(knitr.kable.NA = '')
```


```{r}
theme_update(text = element_text(family = "sans", size=14),
             plot.title = element_text(size=rel(1.2), hjust = 0),
             plot.caption = element_text(size=rel(1.0), hjust = 0, colour="grey10"),
             plot.margin = margin(15, 0, 15, 0))
```



```{r include=FALSE}
# data wrangling
dt <- fread("data-imdb-movies.csv")

# rank by rating (high to low)
dt <- cbind(1:nrow(dt), dt[order(rating, decreasing = T)])
colnames(dt)[1] <- "rank"
dt[, rating_level := ifelse(rating>=7.6, "High", "Low")]
dt$rating_level <- factor(dt$rating_level, levels=c("High", "Low"))

# adjust for inflation
dt$gross_world <- 1.0178^(2019-dt$year) * dt$gross_world
dt$gross_domestic <- 1.0178^(2019-dt$year) * dt$gross_domestic

# correct errors in worldwide gross values
idx <- which(dt$gross_domestic > dt$gross_world)
dt[idx]$gross_world <- dt[idx]$gross_world + dt[idx]$gross_domestic

# impute missing gross box office
dt[is.na(gross_domestic)]$gross_domestic <- median(dt$gross_domestic, na.rm=T)
dt[is.na(gross_world)]$gross_world <- median(dt$gross_world, na.rm=T)

# time elasped since 2010 in years
dt$year_offset <- dt$year - 2010

# convert to factor
dt$year <- factor(dt$year, levels=sort(unique(dt$year)))
dt$genre <- factor(dt$genre, levels=data.table(sort(table(dt$genre), decreasing = T))$V1)

# % of #votes from each demographic group
dt$votesMperc <- dt$votesM / dt$votes * 100
dt$votesFperc <- dt$votesF / dt$votes * 100
dt$votesU18perc <- dt$votesU18 / dt$votes * 100
dt$votes1829perc <- dt$votes1829 / dt$votes * 100
dt$votes3044perc <- dt$votes3044 / dt$votes * 100
dt$votes45Aperc <- dt$votes45A / dt$votes * 100
dt$votesUSperc <- dt$votesUS / dt$votes * 100
dt$votesNonUSperc <- dt$votesNonUS / dt$votes *100

# within demo type perc
dt$Fperc <- dt$votesF / (dt$votesF + dt$votesM) * 100
dt$`<30perc` <- (dt$votesU18+dt$votes1829) / (dt$votesU18+dt$votes1829+dt$votes3044+dt$votes45A) * 100
dt$USperc <- dt$votesUS / dt$votesNonUS * 100

# votes ratios
dt$`F/M` <- dt$votesFperc / dt$votesMperc
dt$`<30/30+` <- (dt$votesU18perc + dt$votes1829perc) / (dt$votes3044perc + dt$votes45Aperc)
dt$`US/nonUS` <- dt$votesUSperc / dt$votesNonUSperc

# votes percentage diff
dt$`A30-U30` <- -(dt$votesU18perc + dt$votes1829perc - dt$votes3044perc - dt$votes45Aperc)
dt$`M-F` <- dt$votesMperc -dt$votesFperc
dt$`nonUS-US` <- dt$votesNonUSperc -dt$votesUSperc

# new data tables showing ratings and votes by gender/age/country in long form
dbasic <- dt[, .(rank, title, year, genre, runtime, gross_world, gross_domestic, director, star)]

dsub <- function(suffix, newcolname, newcolval) {
  d <- dt[, .(rank, rating_level,
              get(paste0("rating",suffix)), 
              get(paste0("votes",suffix)), 
              get(paste0("votes",suffix,"perc")))]
  colnames(d)[2:5] <- c("rating_level", "rating", "votes", "votesperc")
  d[, newcolname] <- newcolval
  d$type <- newcolname
  return(d)
}

dgender <- merge(dbasic, rbind(
    dsub("M", "Gender", "Male"), 
    dsub("F", "Gender", "Female")),
  by="rank")

dage <- merge(dbasic, rbind(
    dsub("U18", "Age", "<18"), 
    dsub("1829", "Age", "18-29"), 
    dsub("3044", "Age", "30-44"), 
    dsub("45A", "Age", "45+")),
  by="rank")

dlocation <- merge(dbasic, rbind(
    dsub("US", "Location", "U.S."), 
    dsub("NonUS", "Location", "Non U.S.")),
  by="rank")

ddemo <- rbind(dgender, dage, dlocation, use.names=FALSE)
colnames(ddemo)[which(colnames(ddemo)=="Gender")] <- "demographic"

# new director and star datasets
get_mean_dt <- function(rowidx=1:nrow(dt), bycol="director") {
  d <- dt[rowidx, .(.N, mean(rank), mean(rating), mean(votes), 
                    mean(gross_world, na.rm=T), mean(gross_domestic, na.rm=T)),
                by=bycol]
  colnames(d) <- c(bycol, "freq", "rank", "rating", "votes", "gross_world", "gross_domestic")
  
  d$freq <- factor(d$freq, levels=sort(unique(d$freq), decreasing=T))
  ord <- d[order(freq, rating)][, get(bycol)]
  d[, eval(bycol) := factor(get(bycol), levels=ord)]
  return(d)
}

ddirector <- get_mean_dt()
ddirector_high <- get_mean_dt(dt$rating_level=="High")
ddirector_low <- get_mean_dt(dt$rating_level=="Low")
dstar <- get_mean_dt(bycol="star")
dstar_high <- get_mean_dt(dt$rating_level=="High", "star")
dstar_low <- get_mean_dt(dt$rating_level=="Low", "star")

dt$director <- factor(dt$director, levels=levels(ddirector$director))
dt$star <- factor(dt$star, levels=levels(dstar$star))
dt[, director_freq:=.N, by=director]
dt[, director_rating:=mean(rating), by=director]
dt[, star_freq:=.N, by=star]
dt[, star_rating:=mean(rating), by=star]

dt_high <- dt[rating_level=="High"]
dt_low <- dt[rating_level=="Low"]
```

Each section corresponds to one interactive graph. They explore associations between the response variable `rating` and different predictors. Each graph is also followed by a foldable *brief discussions* section.

### Rating vs. Votes, Runtime, Domestics Gross, and Worldwide Gross

To begin, Figure 1 shows the correlation between rating and continuous attributes through correlation coefficients and scatter plots. In addition, it compares the attributes' histogram distributions between higher-rated and lower-rated movies.

Since the mean of the rating distribution is about 7.1, Figure 1 uses >7 (i.e. ≥7.1) as the default definition of high ratings. Using the slide bar, we can set different rating thresholds.

<div style="text-align: left; margin-left: 64px; margin-bottom:0px; top: 18px; position: relative">

<p style="margin-bottom:0px">

Rating Level Threshold: [7]{#G1threshold} <br> Low ≤
<input type="range" id="G1slider" name="rating threshold" min="4.5" max="8.5" value="7" step="0.5" onchange="updateG1(this.value);" list="steplist">
<datalist id="steplist" style="margin-bottom:0px"> <option>4.5</option> <option>5.0</option>
<option>5.5</option> <option>6.0</option> <option>6.5</option>
<option>7.0</option> <option>7.5</option> <option>8.0</option>
<option>8.5</option> </datalist> \<High

</p>

</div>

```{r fig.width=8.28, fig.height=4.5}
## Figure 1
cols <- c("rating", "votes", "runtime", "gross_domestic", "gross_world")
G1longdt <- data.table(pivot_longer(data.frame(dt[,..cols]), cols=all_of(c("votes", "runtime", "gross_domestic", "gross_world"))))

G1longdt[name=="votes", value := value/1000]

G1longdt$name <- factor(G1longdt$name, levels=c("votes", "runtime", "gross_domestic", "gross_world"), 
                      labels=c("Total Votes (thousands)", "Runtime (mins)", "Domestic Gross ($millions)", "Worldwide Gross ($millions)"))
G1longdt$rating_level <- "high"
colnames(G1longdt) <- c("rating", "name", "value", "rating level")

cors <- round(cor(dt[,..cols])[2:5,1],3)
G1longdt$name2 <- factor(G1longdt$name, levels=levels(G1longdt$name), labels=paste("Corr:", cors))
G1longdt1 <- G1longdt
G1longdt1$value <- round(G1longdt1$value, 1)

plotG1 <- function(th) {
  
  lvl <- c("high", "low")
  clrs <- c("#F87660", "#00BFC4")
  if (th>= 7.5) {
     lvl <- rev(lvl)
     clrs <- rev(clrs)
  }
  G1longdt$`rating level` <- factor(ifelse(G1longdt$rating>th, "high", "low"), levels=lvl)
  G1longdt1$`rating level` <- G1longdt$`rating level`

  g1 <- ggplot(G1longdt, aes(x=value, fill=`rating level`)) + 
    scale_fill_manual(values=clrs) +
    geom_histogram(bins=23, position="identity", alpha=0.38) +
    facet_wrap(~name, scales="free_x", nrow=1) +
    theme(text = element_text(size = 10), 
          plot.caption = element_text(size = 10)) +
    labs(y=element_blank(), x=element_blank(), fill="Rating Level") + 
    theme_minimal() +
    theme(panel.spacing.y = unit(3,"lines"), 
          axis.ticks.x = element_blank(), 
          axis.text.x = element_blank(), 
          strip.text.x = element_text(size=10))


  g11 <- ggplot(G1longdt1, aes(x=value, y=rating, color=`rating level`)) + 
    scale_color_manual(values=clrs) +
    geom_point(alpha=0.2) +
    facet_wrap(~name2, scales="free_x", nrow=1) +
    theme(text = element_text(size = 10), plot.caption = element_text(size = 10)) +
    geom_abline(slope=0, intercept = th, alpha=0.3, lty=2, lwd=1) +
    labs(title="Movie Rating vs. Continuous Variables", y="Rating", x=element_blank(), 
         color="Rating Level") + 
    theme_minimal() +
    theme(panel.spacing.y = unit(3,"lines"))


  return(subplot(
    with_options(list(digits=2), style(ggplotly(g1, tooltip=c("y","x","fill")), showlegend=F)),
    with_options(list(digits=2), style(ggplotly(g11, tooltip=c("y","x")), showlegend=F)), 
    titleX=T, titleY=T, nrows = 2, margin=0.022, heights=c(0.25, 0.75)
  ) %>%
    layout(
      yaxis2=list(title="Rating"), font=list(size=12),
      annotations=list(
        list(x=0, y=-0.11, showarrow = F, xref='paper', yref='paper', font=list(size=12), text=paste0("Figure 1. The continuous attributes' correlation with rating and distribution by rating level with threshold ", th, ".")),
        list(x=0.998, y=0.722, showarrow = F, xref='paper', yref='paper', font=list(size=12, color="#F87660"), text="high"),
        list(x=0.995, y=0.02, showarrow = F, xref='paper', yref='paper', font=list(size=12, color="#00BFC4"), text="low")
      )
    ) )
}

thresholds <- seq(4.5, 8.5, 0.5)
G1s <- list()
for (i in 1:length(thresholds)) {
  G1s[[i]] <- plotG1(thresholds[i])
}
thresholds <- paste(thresholds, collapse=',')
G1s[[1]]
G1s[[2]]
G1s[[3]]
G1s[[4]]
G1s[[5]]
G1s[[6]]
G1s[[7]]
G1s[[8]]
G1s[[9]]
```


```{=html}
<script>
var thresholds = "`r thresholds`".split(',')
</script>
```
```{js}
const G1s = document.getElementsByClassName("plotly html-widget")
var prev_th = Number(document.getElementById('G1threshold').innerText)
const min_th = Number(thresholds[0])
const n_th = thresholds.length

for (var i = 0; i < n_th; i++) {
  G1s[i].style.display = "none"
}
updateG1(prev_th)

function updateG1(th) {
  document.getElementById('G1threshold').innerText = th;
  G1s[parseInt((prev_th-min_th) / 0.5)].style.display = "none"
  G1s[parseInt((Number(th)-min_th) / 0.5)].style.display = "table"
  prev_th = Number(th)
}
```

<details>
  <summary>Brief discussions</summary>
By both the histograms and the scatter plot, there is a moderate, positive correlation between the number of votes and the rating. There is also a weak, positive correlation between rating and runtime. In comparison, there is not an obvious, significant correlation between rating and the box office variables. The distributions of all four predictors are more or less right-skewed. In all four cases, the most extreme outliers have above-average ratings.

</details>

### Rating vs. Genre and Year

Next, Figure 2 compares the mean rating of movies across different genres, years, as well as genre-year combinations. The genre-year distribution is illustrated through bubble sizes; and the total counts of each genre and year can also be checked by hovering over the squares.

```{r fig.width=5.8, fig.height=4.8}
d <- dt[, .(count = .N, rating = mean(rating)), by=c("genre", "year")]
d <- rbind(dt[, .(genre="(any)", year="(any)", count=0, rating=mean(rating))], d)[-c(1),]

d2 <- rbind(dt[,.(genre="(any)", year="(any)", count=.N, rating=mean(rating))],
            rbind(
  dt[, .(year = "(any)", count = .N, rating=mean(rating)), by="genre"],
  dt[, .(genre = "(any)", count = .N, rating=mean(rating)), by="year"])
  )
d2$rating <- round(d2$rating, 1)

g2 <- ggplot() +
  scale_color_distiller(palette="RdYlBu") +
  geom_point(data=d, aes(year, genre, color=rating, size=count)) +
  geom_point(data=d2, aes(year, genre, color=rating, label=count), size=5, shape=15) +
  geom_text(aes(year, genre, label=rating, fill=count), data=d2, color="grey35", size=3) +
  geom_point(data=d2, aes(year, genre), color="grey70", size=5.5, shape=0, stroke=0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = -0.2)) +
  labs(title="Movie Count and Rating by Genre and Year",
       x="Year\n.", y="Genre", color="Rating", size="Count") +
  scale_x_discrete(limits=levels(d$year)) +
  scale_y_discrete(limits=levels(d$genre))

with_options(list(digits=2), ggplotly(g2)) %>%
  layout(annotations=list(
    x=0, y=-0.173, showarrow = F, xref='paper', yref='paper',
    font=list(size=12), xshift=-18, text="Figure 2. The count (size) and mean rating (color) by genre and year.") 
  ) %>%
  style(hoverinfo = "none", traces = c(3,4))
```

<details>
  <summary>Brief discussions</summary>
The averaged movie ratings in different years are similar, with slightly higher average ratings in years 2014-2017. 

In comparison, there are a lot more deviations across the genre categories. There is only one mystery movie and one scientific fiction movie. Although their ratings are above the average, we should not conclude this as a typical phenomenon. Among groups with a reasonably large size, crime, animation and biography movies tend to be highly rated. In addition, a majority of the drama movies have a good rating, with a few exceptionally low ratings in the second half of the decade. On the other hand, horror movies and comedy movies tend to have a lower rating than average. Comedy movies were very popular but received low ratings in the first demi-decade, which is in contrast to the lower counts but higher ratings in the second demi-decade. 


</details>

### Rating vs. Star and Director

There are 185 unique stars and 200 unique directors. Since there are too many categories, I performed frequency encoding. The star/director frequency represents the number of times a star/director occurs in this dataset.

Figure 3 was created using `plot_ly(type="treemap")`. It explores how movie ratings may correlate with star and director frequencies. Each individual tile represents one movie; and these movie tiles are grouped by the frequency tiles. A tile's colour corresponds to its mean rating. 

To compare the counts and mean ratings of frequency groups, we can either look at the tiles' sizes and colours or hover over the tiles.
We can also click on each individual movie tile to look at its detailed movie information. Click again to go back.


```{r fig.width=8.3}
## Figure 3
image_file <- "./img/scale.png"
txt <- RCurl::base64Encode(readBin(image_file, "raw", file.info(image_file)[1, "size"]), "txt")

plotly_treemap <- function(col="director", p=NA) { 
  if (col == "director") {
    d <- ddirector
    type <- "Director"
    abbr <- "Director"
    col_freq <- "director_freq"
  } else {
    d <- dstar
    type <- "Star"
    abbr <- "Star"
    col <- "star"
    col_freq <- "star_freq"
  }

  lab <- paste("Freq", as.character(levels(d$freq)))
  title <- "" 
  par <- rep(title, times=length(lab))
  nmovies <- dt[,.N, by = col_freq][order(get(col_freq), decreasing=T), N]
  np <- nmovies / as.integer(levels(d$freq))
  txt <- paste0("---\nNo. Movies : ", nmovies, "\nNo. Unique ", type, "s : ", np, rep("\nMean ", times=length(lab)))
  #lab <- c(lab, paste0(type, "Freq : ", dt[,get(col_freq)], '\n"', dt$title, '"'))
  lab <- c(lab, paste0('"', dt$title, '"\n---'))
  par <- c(par, paste("Freq", dt[, get(col_freq)]))
  #val <- c(300, 20, 18, 106, 136)
  #val <- c(val, frequency)
  val <- unlist(dt[, mean(rating), by=col_freq][order(get(col_freq), decreasing=T),2], use.names=F)
  val <- c(val, dt$rating)
  
  txt <- paste0(c(txt, paste0(
    paste(abbr, ":", dt[, get(col)], "\n"),
    paste(type, "Frequency :", dt[,get(col_freq)], "\n"))), 
    paste("Rating :",round(val, 1)))
   
  
  df <- data.frame(x=1:length(val), y=1, col=val)
  
  # construct palette using ggplot
  g <- ggplot(df, aes(x=x, y=y, color=col)) + geom_point(size=1) +
    scale_color_distiller(palette="RdYlBu")
  clr <- ggplot_build(g)$data[[1]]$colour
  
  if (is.na(p)) {
    plot_ly(
      type="treemap",
      labels=lab,
      parents=par,
      marker=list(colors=clr),
      text=txt,
      sort=F,
      domain=list(column=0),
      hovertemplate="%{label}<br>%{text}<extra></extra>",
      textposition="middle left"
    )
  } else {
    add_trace(p,
      type="treemap",
      labels=lab,
      parents=par,
      marker=list(colors=clr),
      text=txt,
      sort=F,
      domain=list(column=1),
      hovertemplate="%{label}<br>%{text}<extra></extra>",
      textposition="middle left"
    )
  }
}

p <- plotly_treemap("star")
plotly_treemap("director", p) %>%
  layout(
    grid=list(columns=2, rows=1), tracegroupgap=0,
    margin=list(l=30, r=65, b=33, t=0),
  annotations=list(list(
    x=0, y=-0.035, xshift=3, showarrow = F, xref='paper', yref='paper',
    font=list(size=12, color="black"), text="Figure 3. Movies and ratings by star and director frequencies. There are 300 movies, 185 stars, and 200 directors."),
    list(x=0.12, y=1, showarrow=F, xref='paper', yref='paper', font=list(size=14, color="black"), text="Movies by Star Frequency"),
    list(x=0.92, y=1, showarrow=F, xref='paper', yref='paper', font=list(size=14, color="black"), text="Movies by Director Frequency")
    ),
  images=list(list(source=paste('data:image/png;base64', txt, sep=','), xref='paper', yref='paper', x=1.025, y=0.66, sizex=0.9, sizey=0.302, layer="above"))
  )

```

<details>
  <summary>Brief discussions</summary>
Based on the size and the hover info of the grouping tiles, over 40% of the movies belong to the *freq 1* group.
Overall, there is an obvious increase in the mean rating as the director frequency increases. In contrast, we do not observe any significant linear relationship between the star frequency and the movie rating. Surprisingly, the group of frequency 8 is the one with the lowest mean rating. This potentially suggests that the movie quality depends more on the experience of the director rather than the star.
<br>
<br>
</details>